Prestige as a Driving Force in Cultural Transmission

### 2022-05-15 ms/PrestigeBias.tex
update

\begin{result}[GBD approximation]
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

@@

% TODO in the following, I would organize differently:
% 1.  remark for E[x'-x] the "drift" (1st order) coefficient in a diffusion equation approximation
% 2. result on effective selection coefficient
% 3. remark for V[x'-x] the "diffusion" (2nd order) coefficient in a diffusion equation approximation
% 4. result of effective population size
% 5. result for fixation prob
% 6. result for fixation time
% 7. result for changing environment
% so I didn't read the rest of the results very carefully, waiting for you to think about the re-arrangement suggestion first.

@@

\begin{result}[Effective population size]\label{res:effective_population}
$N_e=\alpha N + (1-\alpha)$, where $Ne$ is the effective population size of our binary model.
\end{result}

@@

We suspect that when $\beta$ is too large, there won't be many cycles in the simulations. This might happen if either the population reaches a high frequency of the fitter phenotype after only a few cycles, or the fitter phenotype get extinct very quickly. 
For the very high $\beta$ values (0.35 and above), the fixation probability exceeds even the constant environment approximation (which is the upper limit), but it's to be expected, because Kimura's equations are only appropriate for weak fitness (i.e., low selection coefficients, $s$).

@@

\section*{Discussion}
\paragraph{Summary.}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be more likely to be copied due to transmission biases. 
A common one is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be correctly identified.
However, it has been suggested \citep{complexityPaper} that because estimating success is hard, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful by some other trait.

Inspired by a model by \citet{evolutionBook}, we developed a cultural-evolution model with prestige bias, which included both success and influence biases, where the latter is a bias towards role-models with many copiers. We found approximations for the role-model choice process: the generalized binomial distribution that approximates the number of copiers of a single role-model, and the Dirichlet-Multinomial distribution that approximates the distribution of copiers in the entire population in a \textit{rich-getting-richer} type of process. 
We then studied the effects of prestige bias on the fixation probability and time of an invading advantageous cultural phenotype in both a constant and a changing environment.
% TODO what did we find?
% TODO I think you can do a better job summarizing the results - does the approximation work? when doesn't it work? why?

>>>>

\begin{result}[GBD approximation]\label{res:GBD}
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

@@

\begin{remark}[Effective selection coefficient]
In population genetics a selection coefficient is a measure of differences in relative fitness. Selection coefficients are central to the quantitative description of evolution, since fitness differences determine the change in genotype frequencies attributable to selection.
Effective selection coefficient is the equivalent in a model to the selection coefficient in a Wright-Fisher model.
\end{remark}

@@

\begin{remark}
The effective size of a population, $N_e$, determines the rate of change in the composition of a population caused by genetic drift, which is the random sampling of genetic variants in a finite population. Ne is crucial in determining the level of variability in a population, and the effectiveness of selection relative to drift.
\end{remark}

\begin{result}[Effective population size]\label{res:effective_population}
$N_e=\alpha N + (1-\alpha)$, where $Ne$ is the effective population size of the binary model.
\end{result}

@@

We suspect that when $1-\beta$ is too large, there won't be many cycles in the simulations. This might happen if either the population reaches a high frequency of the fitter phenotype after only a few cycles, or the fitter phenotype get extinct very quickly. 
For the very high $1-\beta$ values (0.35 and above), the fixation probability exceeds even the constant environment approximation (which is the upper limit), but it's to be expected, because Kimura's equations are only appropriate for weak fitness (i.e., low selection coefficients, $s$).

@@

\section*{Discussion}
\paragraph{Summary.}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be more likely to be copied due to transmission biases. 
A common one is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested \citep{complexityPaper} that because estimating success is hard, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful, portrayed by other traits or means.

Inspired by a model by \citet{evolutionBook}, we developed a cultural-evolution model with prestige bias, which included both success and influence biases, where the latter is a bias towards role-models with many copiers.
As shown in \cref{fig:influence_advantage}, given the option to choose the ratio of success and influence, copiers will minimize their estimation error, hence will copy the traits that are closer to the ideal phenotype. We see that throughout the choosing process, the estimation error decreases, and the success bias rate also decreases. This suggests that the later a copier gets to choose, the less he should rely on his own estimation, and more on previous choices.
In addition, we see that the process of choosing itself is necessary, and when using a constant ratio ($\alpha$), the mean estimation error in the population is greater by a factor of $10$, even when using the $\alpha$ the population converged to when being able to choose it (\cref{fig:const_influence}).
After we showing that choosing one's success bias weight is beneficial to the copier, we continued to create our models- continuous and dichotomous, that incorporated both success and influence. However, running simulations for a nested iterative model is costly, and analyzing such model mathematically is very hard. We therefor found approximations for the role-model choice process: the generalized binomial distribution (\cref{res:GBD}) and the Dirichlet-Multinomial distribution (\cref{cor:dirichlet}).
To use approximate our model with the GBD distribution, we must assume homogeneous estimation error, $e$, and homogeneous success bias weight, $\alpha$. Both assumptions are necessary to this approximation, because it requires all copiers to be identical (order of copiers must not be a factor).
For the DM approximation we can relax the assumption for $\alpha$, and allow a varying $\alpha_j$ for each role-model $j$. This means that a copier is still not able to choose a success bias weight for itself, and instead it is a characteristic of the role-model only.
Using the DM approximation, because it is easier to analyze, we found equivalents for Kimura's equations, for our dichotomous model.
Following \cref{res:selection_coef,res:effective_population} we found an approximation for the fixation probability of the ideal trait, and an approximation for the time of fixation. These approximations hold only when the equivalent for the selection coefficient, in our model $1-\beta(A)$, is low (0.4 or lower). We also discovered that changing the success bias weight doesn't hurt the goodness of the approximations, and only affects the effective population size $N_e$.
Once we found the equations for a constant environment, we moved on to a cyclic changing environment.
Based on \cref{res:ch_expected,res:ch_var}, we found the approximations for the fixation probability and time to fixation.
We discovered that as in the constant environment, the effective population size doesn't affect the goodness of fit of the approximation. The success coefficient however, must be even lower than before for the approximation to fit. When the success coefficient is higher than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation. We believe the reason is the structure of the cycle.
Our proof and approximation in the changing environment are for a large amount of cycles, and when the success coefficient is too high, there might be very few cycles. Either the ideal trait is copied by enough copiers so that the influence is sufficient to negate the success bias when the cycle changes (and the beneficial trait becomes the maladaptive), or the opposite happens, and the ideal trait gets extinct before there were enough copiers that copied it.
We then tried to change the ratio between the number of cycles where $\hat{A}$ is beneficial and maladaptive. We showed that the approximation fits well regardless of the ratio, but when the ratio of adaptive generations to maladaptive ones is very high, it is very similar to a constant environment model.
